---
title: "Project"
author: "Fiona Zhang"
date: "2024-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
input=read.table("Data2share.txt",sep="\t",header=T)

input <- input |>
  mutate(ever_symp = across(starts_with("symptom")) %>% rowSums > 0,
         ever_pos = across(starts_with("compar")) %>% rowSums(., na.rm = TRUE) > 0)
```

```{r}
long_df <- input |>
  select(-eligible) |>
  pivot_longer(
    cols = -newid, 
    names_to = c(".value", "day"), 
    names_pattern = "(.*)\\.(.*)"
  ) |>
  mutate(day = as.numeric(day))
```

```{r}
set.seed(1)
samp <- long_df |>
  distinct(newid) |>
  slice_sample(n = 1000)
long_df |>
  inner_join(samp, by = "newid") |>
  inner_join(input |> select(newid, ever_symp, ever_pos), by = "newid") |>
ggplot(aes(x = day, y = antigen_result, group = newid, color = ever_symp)) +
  geom_line(alpha = 0.5) +
  facet_wrap(~ever_symp)

long_df |>
  inner_join(input |> select(newid, ever_symp, ever_pos), by = "newid") |>
  filter(ever_pos) |>
  group_by(day, ever_symp) |>
  summarise(m = mean(antigen_result, na.rm = TRUE)) |>
ggplot(aes(x = day, y = m, color = ever_symp)) +
  geom_line() +
  facet_wrap(~ever_symp)
```

```{r}
long_df |>
  group_by(newid)
```

```{r}
set.seed(1)
samp <- long_df |>
  distinct(newid) |>
  slice_sample(n = 1000)

true_data <- long_df %>%
 # inner_join(samp, by = "newid") |>
  inner_join(input |> select(newid, ever_symp, ever_pos), by = "newid") %>%
  filter(ever_symp == TRUE, ever_pos == TRUE)

model_true <- lm(antigen_result ~ splines::ns(day, 4), data = true_data) 

true_data |>
  drop_na(antigen_result, day) |>
  mutate(yhat = fitted(model_true)) |>
  distinct(day, yhat) |>
  ggplot(aes(x = day, y = yhat)) + 
  geom_point() + 
  geom_line() + 
  labs(
    title = "Predicted Antigen result from day of test among symptomatic participants",
    subtitle = "Linear Model with Spline")

# true_data |>
#   filter(ever_pos) |>
#   group_by(day) |>
#   summarise(m = mean(antigen_result, na.rm = TRUE)) |>
# 
#   ggplot(aes(x = day, y = m)) +
#   geom_line() 
# ggplot(true_data, aes(x = day, y = antigen_result)) + 
#   geom_point() +
#   stat_smooth(method = "lm") + 
#   labs(title = "Predicted Antigen result from day of test among symptomatic participants",
#        subtitle = "Linear Model (no spline)")


false_data <- long_df %>%
 # inner_join(samp, by = "newid") |>
  inner_join(input |> select(newid, ever_symp, ever_pos), by = "newid") %>%
  filter(ever_symp == FALSE, ever_pos == TRUE)

model_false <- lm(antigen_result ~ splines::ns(day, 4), data = false_data) 

sym_data <- true_data |>
  drop_na(antigen_result, day) |>
  mutate(yhat = fitted(model_true))

p0 <- false_data |>
  drop_na(antigen_result, day) |>
  mutate(yhat = fitted(model_false)) |>
  ggplot(aes(x = day, y = yhat)) + 
  geom_point() + 
  geom_line() + 
  geom_point(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue") + 
  geom_line(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue")  +
  labs(title = "without a lag")

p1 <- false_data |>
  drop_na(antigen_result, day) |>
  mutate(yhat = fitted(model_false),
         day = day + 1) |>
  ggplot(aes(x = day, y = yhat)) + 
  geom_point() + 
  geom_line() + 
  geom_point(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue") + 
  geom_line(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue")  +
  labs(title = "1 day lag")

p2 <- false_data |>
  drop_na(antigen_result, day) |>
  mutate(yhat = fitted(model_false),
         day = day + 2) |>
  ggplot(aes(x = day, y = yhat)) + 
  geom_point() + 
  geom_line() + 
  geom_point(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue") + 
  geom_line(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue")  +
  labs(title = "2 day lag")

p3 <- false_data |>
  drop_na(antigen_result, day) |>
  mutate(yhat = fitted(model_false),
         day = day + 3) |>
  ggplot(aes(x = day, y = yhat)) + 
  geom_point() + 
  geom_line() + 
  geom_point(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue") + 
  geom_line(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue")  +
  labs(title = "3 day lag")

p4 <- false_data |>
  drop_na(antigen_result, day) |>
  mutate(yhat = fitted(model_false),
         day = day + 4) |>
  ggplot(aes(x = day, y = yhat)) + 
  geom_point() + 
  geom_line() + 
  geom_point(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue") + 
  geom_line(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue")  +
  labs(title = "4 day lag")


p5 <- false_data |>
  drop_na(antigen_result, day) |>
  mutate(yhat = fitted(model_false),
         day = day + 5) |>
  ggplot(aes(x = day, y = yhat)) + 
  geom_point() + 
  geom_line() + 
  geom_point(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue") + 
  geom_line(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue")  +
  labs(title = "5 day lag")

p6 <- false_data |>
  drop_na(antigen_result, day) |>
  mutate(yhat = fitted(model_false),
         day = day + 6) |>
  ggplot(aes(x = day, y = yhat)) + 
  geom_point() + 
  geom_line() + 
  geom_point(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue") + 
  geom_line(data = sym_data, aes(x = day, y = yhat), color = "cornflower blue")  +
  labs(title = "6 day lag")
```
We can use a change point detection model to try to determine what we should be adding to the asymptomatic days to get them to match the shape of the symptomatic days. 

* Treat the symptomatic curve as the "truth" and then shift the asymptomatic curve 0.5 days at a time to see if you can find the optimal shift ("optimal" can be defined as the shift that minimizes the MSE between the symptomatic curve and the shifted asymptomatic curve)

```{r}
library(patchwork)
p0 + p1 + p2 + p3 + p4 + p5 + p6

```


```{r}
true_data <- true_data|>
  mutate(antigen_1daylag = lag(antigen_result)) |>
  mutate(antigen_2daylag = lag(antigenyes),
         )
model <- lm(day ~ antigen_result + antigen_1daylag + antigen_2daylag, data = true_data)

true_data |>
  filter(!is.na(antigen_1daylag), !is.na(antigen_2daylag), !is.na(antigen_result), !is.na(day)) |>
  ggplot(aes(x = day, antigen_result)) +
  geom_point() + 
  geom_line(aes(x = fitted(model)))

true_data <- true_data|>
  drop_na(antigenyes) |>
  mutate(antigenyes_hat = fitted(model_yes))
model_bye <- lm(antigenbye ~ splines::ns(day + antigenyes_hat, 4), data = true_data)
true_data <- true_data|>
  drop_na(antigenbye) |>
  mutate(antigenbye_hat = fitted(model_bye))
model_day <- lm(day ~ splines::ns(antigen_result + antigenyes_hat + antigenbye_hat, 4), data = true_data)
true_data <- true_data |>
  mutate(dayhat = fitted(model_day))
ggplot(true_data, aes(x = dayhat, y = antigen_result + antigenyes_hat + antigenbye_hat)) + 
  geom_point() +
  geom_line()
```

